---
title: 'Data Analysis'
author: 
output: 
  github_document:
    toc: yes
  #pdf_document:
  #  dev: cairo_pdf
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = "center",
                      fig.path=('../../report/figures/'))
library(printr)
options(digits = 2)
library(rmarkdown)
```

```{r "env", message=FALSE, warning=F}
rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
source(here::here("src/lib/functions.R"))
#renv::snapshot()
```

```{r "get data"}
load(here("data/intermediate/data_for_analyses.RData"))
```

## Correlation Matrix all variables
```{r "correlation-matrix", echo=F, message=T, warning=T, fig.width=8, fig.height=8}
d %>%
  dplyr::select(`Preference for Narrative Complexity` = PNC,
         `Need for Cognition` = NFC,
         `Tolerance towards Ambiguity` = TOA, 
         `Left-Right Ideological Self-Placement` = E2, 
         `Progressive-Conservative Ideological Self-Placement` = E2_2,
         `Exposure to Complex Media`  = MCE,
         Gender = sex,
         Age = age, 
         Education = education, 
         Urbanization = urbanisation) %>%
  ggcorrmat(
  type = "robust", # correlation method
  cor.vars = c(everything()), # a range of variables can be selected
  colors = c("violet", "white", "seagreen"),
  title = "Correlalogram for Variables under Study"
)

```


## H1: Individuals with a higher PNC are more likely to exhibit greater tolerance for ambiguity (H1a), demonstrate a higher need for cognition (H1b)
### DV = PNC

PNC is predicted by....
- TOA (positive)
- NFC (positive)
- age
- sex
- education


1. Significant effect of XXX = more TOA leads to more XXX

```{r "testing-h1", fig.width=18, fig.height=18}
source(here("src/analysis/h1.R"))

h1
p1
```


## H2: Individuals with a lower PNC are more likely to have a more right-wing political orientation (H2a) and will vote for conservative political parties (H2b).

  
```{r "testing-h2", fig.width=18, fig.height=18}
source(here("src/analysis/h2.R"))

h2_1
h2_2
p2
```

